---
title: "HW10"
author: "Annie"
date: "2025-04-26"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Use the “dominant” copepod species data (from HW1).
Apply PCA or MDS using matrix algebra to extract major gradients of the dominant species. 
Make a bi plot to show the relationships among species and sites. 
Then, check your results with those from build in functions of PCA, MDS, and NMDS.
Hint: use SVD or eigen analysis

Find dominant species
```{r}
copepod_data <- read.table("copepod_composition.txt", header = TRUE)
cop_density <- read.table("cop_density.txt", header = FALSE)
species_list <- read.table("copepodSPlist.txt", fill=T, sep = "\n")

row.names(copepod_data) <- species_list$V1


# Prepare an empty matrix to store density data for all species
all_species_density <- matrix(0, nrow = nrow(copepod_data), ncol = ncol(copepod_data))
row.names(all_species_density) <- row.names(copepod_data)
colnames(all_species_density) <- colnames(copepod_data)

# Loop through each species (row) and each station/sample (column) to calculate density
for (i in 1:nrow(copepod_data)) {
  for (j in 1:ncol(copepod_data)) {
    # Species proportion × corresponding total density at that station
    all_species_density[i, j] <- copepod_data[i, j] / 100 * cop_density[j, 1]
  }
}

# Identify dominant species
# A dominant species is defined as one that makes up ≥5% of the total composition at any cruise-station
threshold <- 5  # Threshold in percent

# Find species that exceed the threshold at least once across all stations
dominant_species <- rownames(copepod_data)[apply(copepod_data, 1, function(x) any(x >= threshold))]

# Display the dominant species
print(dominant_species)
```

Extract density info of dominant species
```{r}
dom_data <- all_species_density[dominant_species, ]
```

Apply PCA or MDS using matrix algebra to extract major gradients of the dominant species. 
```{r}
# Center the data (subtract the mean, no scaling)
#center = TRUE：每一個物種變數都減去自己的平均值（變成平均為 0）。
#scale = FALSE：不要除以標準差（不做 Z-score 標準化）。
dom_data_centered <- scale(t(dom_data), center = TRUE, scale = FALSE)


# Perform SVD decomposition
svd_result <- svd(dom_data_centered)

# Get site scores (projection onto the lower-dimensional space)
site_scores <- svd_result$u %*% diag(svd_result$d)

# Get species loadings (species directions)
species_loadings <- svd_result$v

```

Plot the Biplot
```{r}
library(ggplot2)
library(ggrepel)

# 計算 scaling factor 讓物種箭頭長度差不多
scaling_factor <- max(abs(site_scores)) / max(abs(species_loadings))

# 把資料整理成 dataframe 方便 ggplot
site_df <- data.frame(PC1 = site_scores[,1],
                      PC2 = site_scores[,2],
                      Site = rownames(dom_data_centered))

species_df <- data.frame(PC1 = species_loadings[,1] * scaling_factor,
                         PC2 = species_loadings[,2] * scaling_factor,
                         Species = rownames(dom_data))

ggplot() +
  geom_point(data = site_df, aes(x = PC1, y = PC2), color = "blue", size = 2) +
  geom_text_repel(data = site_df, aes(x = PC1, y = PC2, label = Site),
                  color = "blue", size = 3, max.overlaps = 100) +
  
  geom_segment(data = species_df, aes(x = 0, y = 0, xend = PC1, yend = PC2),
               arrow = arrow(length = unit(0.2, "cm")), color = "red") +
  geom_text_repel(data = species_df, aes(x = PC1, y = PC2, label = Species),
                  color = "red", size = 3, max.overlaps = 100) +
  
  labs(title = "PCA Biplot (SVD)",
       x = "PC1", y = "PC2") +
  theme_minimal()

```
check your results with those from build in functions of PCA, MDS, and NMDS.

Built-in PCA
```{r}
pca_builtin <- prcomp(t(dom_data), center = TRUE, scale. = FALSE)  

# get site scores (PC1、PC2)
site_scores_pca <- pca_builtin$x[, 1:2]

# get species loadings (PC1、PC2)
species_loadings_pca <- pca_builtin$rotation[, 1:2]

# PCA biplot
site_df_pca <- data.frame(PC1 = site_scores_pca[,1],
                          PC2 = site_scores_pca[,2],
                          Site = rownames(site_scores_pca))

species_df_pca <- data.frame(PC1 = species_loadings_pca[,1] * max(abs(site_scores_pca)) / max(abs(species_loadings_pca)),
                             PC2 = species_loadings_pca[,2] * max(abs(site_scores_pca)) / max(abs(species_loadings_pca)),
                             Species = rownames(species_loadings_pca))

ggplot() +
  geom_point(data = site_df_pca, aes(x = PC1, y = PC2), color = "blue", size = 2) +
  geom_text_repel(data = site_df_pca, aes(x = PC1, y = PC2, label = Site),
                  color = "blue", size = 3, max.overlaps = 100) +
  
  geom_segment(data = species_df_pca, aes(x = 0, y = 0, xend = PC1, yend = PC2),
               arrow = arrow(length = unit(0.2, "cm")), color = "red") +
  geom_text_repel(data = species_df_pca, aes(x = PC1, y = PC2, label = Species),
                  color = "red", size = 3, max.overlaps = 100) +
  
  labs(title = "PCA Biplot (Built-in)",
       x = "PC1", y = "PC2") +
  theme_minimal()

```


MDS
```{r}
# MDS (classical multidimensional scaling)
# 用歐氏距離（Euclidean distance）
distance_matrix <- dist(t(dom_data))  
mds_result <- cmdscale(distance_matrix, k = 2)

# 整理成 dataframe
mds_df <- data.frame(MDS1 = mds_result[,1],
                     MDS2 = mds_result[,2],
                     Site = rownames(mds_result))

# 畫 MDS 散點圖
ggplot(mds_df, aes(x = MDS1, y = MDS2, label = Site)) +
  geom_point(color = "blue", size = 2) +
  geom_text_repel(color = "blue", size = 3, max.overlaps = 100) +
  labs(title = "MDS", x = "MDS1", y = "MDS2") +
  theme_minimal()
```

nMDS
```{r}

library(vegan)

# NMDS (non-metric multidimensional scaling)
# 注意：metaMDS 會自己標準化資料
nmds_result <- metaMDS(t(dom_data), k = 2, trymax = 100)

# 整理成 dataframe
nmds_df <- data.frame(NMDS1 = nmds_result$points[,1],
                      NMDS2 = nmds_result$points[,2],
                      Site = rownames(nmds_result$points))

# 畫 NMDS 散點圖
ggplot(nmds_df, aes(x = NMDS1, y = NMDS2, label = Site)) +
  geom_point(color = "brown", size = 2) +
  geom_text_repel(color = "brown", size = 3, max.overlaps = 100) +
  labs(title = "NMDS (metaMDS)", x = "NMDS1", y = "NMDS2") +
  theme_minimal()



```
比較自製 PCA site scores 和 prcomp() site scores
```{r}

cor(site_scores[, 1:2], site_scores_pca)  # 看相似度，通常 >0.9 很合理

```
[1,]自己算的PC1
[2,]自己算的PC2
自己算的 PC1 跟 prcomp的 PC1 的相關係數 = 1 → 一樣
自己算的 PC2 跟 prcomp的 PC2 的相關係數 = 1 → 也是一樣

其他交叉的 (PC1 vs PC2、PC2 vs PC1) 都是超接近 0 → 完全不相關，正確！


