---
title: "HW2"
author: "Annie"
date: "2025-02-27"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Homework2

1a. Generate 10000 random numbers from Gaussian distribution with mean=20 and variance =10, and plot the distribution.
```{r}

# Generate standard normal variables (mean = 0, sd = 1)
standard_normal <- rnorm(10000)

# Transform to have mean = 20 and variance = 10
gaussian_numbers <- standard_normal * sqrt(10) + 20  

# Plot histogram
hist(gaussian_numbers, breaks = 50, col = "skyblue", 
     main = "Gaussian Distribution", xlab = "Value", probability = TRUE)

#probability: Transform the Y-axis into probability density, ensuring that the total area of the histogram equals 1.
#probability=TRUE: Density; probability=FALSE: frequency
```
1b. Generate 10000 random numbers from Binomial distribution with p=0.5 and n=40, and plot the distribution.

Compare the distribution of 1a and 1b, what do you find? (hint: you can use “rand” and “randn” in matlab.
Or, “runif” and “rnorm” in R)

Ans:
The Gaussian distribution is continuous and has a smoother shape.
The binomial distribution is discrete and may exhibit more obvious peaks.

```{r}
binomial_numbers <- rbinom(10000, size = 40, prob = 0.5)#prob= 0.5: probability of success per trial is 0.5(like tossing coins)
hist(binomial_numbers, breaks = 50, col = "lightcoral", main = "Binomial Distribution", xlab = "Value", probability = TRUE)
```
Make a program that can select our candidates for presentation next week. 
This program should select randomly but avoid selecting the numbers that had been selected before. (hint: uniform distribution, loop, and if statements.) 
```{r}
# Define total students
total_students <- 1:34
students_per_class <- 2
file_name <- "selected_students.txt"  # File to store selected students

# Function to load previously selected students
load_selected_students <- function() {
  if (file.exists(file_name)) {
    selected <- scan(file_name, quiet = TRUE)
    return(selected)
  } else {
    return(c())  # Return empty if file doesn't exist
  }
}

# Function to save selected students
save_selected_students <- function(selected_students) {
  write(selected_students, file = file_name, ncolumns = length(selected_students))
}

# Function to select students
select_students <- function() {
  selected_students <- load_selected_students()  # Load previous selections
  
  # If all students have been selected, stop selecting
  if (length(selected_students) == length(total_students)) {
    cat("All students have been selected.\n")
    return(FALSE)  # Stop selection
  }
  
  # Get remaining students
  remaining_students <- setdiff(total_students, selected_students)
  
  # Shuffle remaining students
  random_index <- order(runif(length(remaining_students)))  # Random order
  remaining_students <- remaining_students[random_index]
  
  # Select the first `students_per_class` students
  selected <- remaining_students[1:students_per_class]
  
  # Update and save selected students
  selected_students <- c(selected_students, selected)
  save_selected_students(selected_students)
  
  cat("Selected students:", selected, "\n")
  
  return(TRUE)  # Continue selection
}

# Run the selection
select_students()



```


