---
title: "HW3"
author: "Annie"
date: "2025-03-06"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

NOTE: You are NOT allowed to use the built-in bootstrap or sampling functions in Matlab or R. 
You can use some basic functions such as find, mean, median, var, sum, max, min, ceil, floor, round, sort, length, size, plot, hist and all matrix manipulation functions (e.g. reshape, ones, zeros). 
You can use “print” to save figures. [hint: use “rand” to generate the bootstrapped samples. When doing regression, sample with pairs (i.e. (X1,Y1), (X1,Y1),(X3,Y3),……]

1. Compute the mean and SE(mean) for the fish and copepod densities (assuming all data points are independent) respectively, using both normal theory and non-parametric bootstrap. 
Plot the histogram of bootstrapped means with bootstrap 1000 times. (Normal theory simply means using the standard formula to estimate SE.) 

read data
```{r}
library(readxl)

# Define the file path (modify this to the actual location of your file)
#file_path <- "C:/Users/User/Desktop/Master 1.2/R/HW/enviANDdensity.xls"

# Read the Excel file
data <- read_excel("C:/Users/User/Desktop/Master 1.2/R/HW/enviANDdensity.xls")

# Extract the required columns and rename them for convenience
fish <- data[["FishDensity (ind./1000m3)"]]
copepod <- data[["CopepodDensity ind./m3"]]

```

using normal theory to compute the SE for fish & copepod densities
```{r}
mean_fish <- mean(fish)
se_fish <- sd(fish) / sqrt(length(fish))

mean_copepod <- mean(copepod)
se_copepod <- sd(copepod) / sqrt(length(copepod)) #SE = SD / √n
```

shows the result
```{r}
print(paste("Normal theory estimates:"))
print(paste("Fish: Mean =", mean_fish, "SE =", se_fish))
print(paste("Copepod: Mean =", mean_copepod, "SE =", se_copepod))
```
using bootstrap to compute the SE for fish & copepod densities
```{r}
B <- 999  # Perform 999 bootstrap resamplings  
n <- length(fish)  # Number of observations in the original dataset  

# Pre-allocate space for storing bootstrap means  
boot_means_fish <- numeric(B + 1)  
boot_means_copepod <- numeric(B + 1)  

# The first sample uses the mean of the original dataset  
boot_means_fish[1] <- mean(fish)  
boot_means_copepod[1] <- mean(copepod)  

# Perform 999 bootstrap resamplings  
for (i in 2:(B + 1)) {  
  # Generate random indices using runif() instead of sample()  
  # runif(n, min = 1, max = n + 1) generates n random numbers in [1, n]  
  # floor() converts them to integers to use as indices  
  random_indices <- floor(runif(n, min = 1, max = n + 1))  
  boot_means_fish[i] <- mean(fish[random_indices])  # Compute mean of the resampled fish data  
  boot_means_copepod[i] <- mean(copepod[random_indices])  # Compute mean of the resampled copepod data  
}  

# Compute standard errors (SE)  
se_boot_fish <- sd(boot_means_fish)  
se_boot_copepod <- sd(boot_means_copepod)  

# Print results  
print("Bootstrap estimates:")  
print(paste("Fish: SE =", se_boot_fish))  
print(paste("Copepod: SE =", se_boot_copepod))  


```
Plot the histogram of bootstrapped means 
```{r}
par(mfrow=c(1,2))  # Set up a 1x2 plotting area


hist(boot_means_fish, main="Bootstrap Means of Fish", 
     xlab="Mean Fish Density", col="lightblue", breaks=30)
abline(v=mean_fish, col="red", lwd=2)

hist(boot_means_copepod, main="Bootstrap Means of Copepod", 
     xlab="Mean Copepod Density", col="lightgreen", breaks=30)
abline(v=mean_copepod, col="red", lwd=2)

```

2. Compute the median and bootstrapped SE(median) for the fish and copepod densities. 
Plot the histogram of bootstrapped medians with bootstrap 1000 times. 

```{r}
# Compute the median of the original data
median_fish <- median(fish)
median_copepod <- median(copepod)

# Print original median values
print("Original Median Values:")
print(paste("Fish: Median =", median_fish))
print(paste("Copepod: Median =", median_copepod))
```
```{r}
# Set bootstrap parameters  
B <- 999  # Perform 999 bootstrap resamplings  
n <- length(fish)  # Number of observations in the original dataset  

# Create vectors to store bootstrap results  
boot_medians_fish <- numeric(B + 1)  
boot_medians_copepod <- numeric(B + 1)  

# The first sample uses the median of the original dataset  
boot_medians_fish[1] <- median_fish  
boot_medians_copepod[1] <- median_copepod  

# Perform 999 bootstrap resamplings  
for (i in 2:(B + 1)) {  
  # Generate random indices using runif() instead of sample()  
  # runif(n, min = 1, max = n + 1) generates n random numbers in [1, n]  
  # floor() converts them to integers to use as indices  
  random_indices <- floor(runif(n, min = 1, max = n + 1))  
  boot_medians_fish[i] <- median(fish[random_indices])  # Compute median for resampled fish data  
  boot_medians_copepod[i] <- median(copepod[random_indices])  # Compute median for resampled copepod data  
}  

# Compute bootstrap standard errors (SE)  
se_boot_median_fish <- sd(boot_medians_fish)  
se_boot_median_copepod <- sd(boot_medians_copepod)  

# Print the results  
print("Bootstrap estimates for Median:")  
print(paste("Fish: Median =", median_fish, "SE =", se_boot_median_fish))  
print(paste("Copepod: Median =", median_copepod, "SE =", se_boot_median_copepod))  



```
Plot histograms of bootstrapped medians
```{r}
par(mfrow=c(1,2))  # Set up a 1x2 plotting area

# Histogram for fish bootstrap medians
hist(boot_medians_fish, main="Bootstrap Medians of Fish", 
     xlab="Median Fish Density", col="lightblue", breaks=30)
abline(v=median_fish, col="red", lwd=2)

# Histogram for copepod bootstrap medians
hist(boot_medians_copepod, main="Bootstrap Medians of Copepod", 
     xlab="Median Copepod Density", col="lightgreen", breaks=30)
abline(v=median_copepod, col="red", lwd=2)
```

3. Plot fish (dependent) v.s copepod (independent) and the regression line. 
Compute the regression coefficients for fish = b0+b1*copepod and bootstrapped SE(b0) and SE(b1). 
Plot the histogram of bootstrapped b0 and b1 with bootstrap 1000 times. 
```{r}
# Compute linear regression coefficients
model <- lm(fish ~ copepod)  #lm: linear model, with fish as the dependent variable and copepod as the independent variable
b0 <- coef(model)[1]  # Extract the intercept (b0) from the regression model
b1 <- coef(model)[2]  # Extract the slope (b1) from the regression model


print("Original Regression Coefficients:")
print(paste("b0 =", b0))
print(paste("b1 =", b1))
```
bootstrapped SE(b0) and SE(b1)
```{r}
# Bootstrap settings
B <- 999  # Perform 999 bootstrap resamplings
n <- length(fish)  # Number of observations in the dataset

boot_b0 <- numeric(B + 1)  # Initialize vector to store intercept estimates
boot_b1 <- numeric(B + 1)  # Initialize vector to store slope estimates

# Fit the original linear model
original_model <- lm(fish ~ copepod)
boot_b0[1] <- coef(original_model)[1]  # Store the intercept of the original model
boot_b1[1] <- coef(original_model)[2]  # Store the slope of the original model

# Bootstrap loop without using sample()
for (i in 2:(B + 1)) {
  # Create a vector of random indices using a uniform distribution
  resample_indices <- floor(runif(n, min = 1, max = n + 1))  # Generate random indices between 1 and n
  
  boot_model <- lm(fish[resample_indices] ~ copepod[resample_indices])  # Fit model on bootstrapped data
  boot_b0[i] <- coef(boot_model)[1]  # Store intercept for bootstrapped model
  boot_b1[i] <- coef(boot_model)[2]  # Store slope for bootstrapped model
}

# Compute bootstrap standard errors
se_boot_b0 <- sd(boot_b0)  # Standard error of the intercept
se_boot_b1 <- sd(boot_b1)  # Standard error of the slope

# Print results
print("Bootstrap estimates for Regression Coefficients:")
print(paste("b0: SE =", se_boot_b0)) 
print(paste("b1: SE =", se_boot_b1)) 


```


```{r}
# Plot regression line
plot(copepod, fish, main = "Regression: Fish vs. Copepod", 
     xlab = "Copepod Density", ylab = "Fish Density", col = "brown", pch = 16)
abline(b0, b1, col = "skyblue", lwd = 2)

# Plot histograms of bootstrapped coefficients
par(mfrow = c(1, 2))

hist(boot_b0, main = "Bootstrap Distribution of b0", 
     xlab = "b0", col = "lightblue", breaks = 30)
abline(v = b0, col = "red", lwd = 2)

hist(boot_b1, main = "Bootstrap Distribution of b1", 
     xlab = "b1", col = "lightgreen", breaks = 30)
abline(v = b1, col = "red", lwd = 2)

```

1. Compute the mean and stand error of the mean for the fish and copepod density (all data points) respectively using Jackknife. 
Plot the histogram of Jackknife means.
```{r}
# Jackknife function to compute mean and standard error
jackknife_mean <- function(data) {
  
  n <- length(data)  # Get the number of data points
  jackknife_means <- numeric(n)  # Create a numeric vector to store the mean after removing each data point.
  
# Loop through each data point, leaving one out at a time
  for (i in 1:n) {
    jackknife_means[i] <- mean(data[-i])  # Compute mean after leaving out the i-th data point
  }
  
  jackknife_mean <- mean(jackknife_means)  # Compute the mean of all leave-one-out means
  jackknife_se <- sqrt((n - 1) / n * sum((jackknife_means - jackknife_mean)^2))  # Compute Jackknife standard error (SE)
  return(list(means = jackknife_means, mean = jackknife_mean, se = jackknife_se))
}

# Calculate for fish and copepod densities
fish_jackknife <- jackknife_mean(fish)
copepod_jackknife <- jackknife_mean(copepod)

# Print results
print(paste("Fish: Jackknife Mean =", round(fish_jackknife$mean, 2), "Jackknife SE =", round(fish_jackknife$se, 2)))
print(paste("Copepod: Jackknife Mean =", round(copepod_jackknife$mean, 2), "Jackknife SE =", round(copepod_jackknife$se, 2)))

# Plot histograms of Jackknife means
par(mfrow=c(1,2))  # Set up a 1x2 plotting area
hist(fish_jackknife$means, main="Jackknife Means of Fish", xlab="Mean Fish Density", col="lightblue", breaks=30)
hist(copepod_jackknife$means, main="Jackknife Means of Copepod", xlab="Mean Copepod Density", col="lightgreen", breaks=30)


```
2. Compute the regression coefficients for fish = b0+b1*copepod and Jackknife SE of b0 and b1.
Plot the histogram of Jackknife b0 and b1.

```{r}
# Jackknife function to compute regression coefficients (b0 and b1)

jackknife_regression <- function(fish, copepod) {
  
  n <- length(fish) # Get the number of data points
  jack_b0 <- numeric(n)  # Initialize vector to store leave-one-out b0 (intercept)
  jack_b1 <- numeric(n)  # Initialize vector to store leave-one-out b1 (slope) 
  
  for (i in 1:n) {
    # Create the jackknife sample (leave out the i-th data point)
    jackknife_fish <- fish[-i] # Remove the i-th fish data point
    jackknife_copepod <- copepod[-i] # Remove the corresponding copepod data point
    
    # Fit the linear model on the remaining data
    model <- lm(jackknife_fish ~ jackknife_copepod) # Perform linear regression
    jack_b0[i] <- coef(model)[1]  # Store b0 (intercept)
    jack_b1[i] <- coef(model)[2]  # Store b1 (slope)
  }
  
  jackknife_b0 <- mean(jack_b0)  # Jackknife estimate of b0
  jackknife_b1 <- mean(jack_b1)  # Jackknife estimate of b1
  
  se_b0 <- sqrt((n - 1) / n * sum((jack_b0 - jackknife_b0)^2))  # Jackknife SE for b0
  se_b1 <- sqrt((n - 1) / n * sum((jack_b1 - jackknife_b1)^2))  # Jackknife SE for b1
  
  return(list(b0 = jackknife_b0, b1 = jackknife_b1, se_b0 = se_b0, se_b1 = se_b1, jack_b0 = jack_b0, jack_b1 = jack_b1))
}

# Calculate for fish and copepod densities
regression_jackknife <- jackknife_regression(fish, copepod)

# Print results
print(paste("Jackknife b0 =", round(regression_jackknife$b0, 2), "SE(b0) =", round(regression_jackknife$se_b0, 2)))
print(paste("Jackknife b1 =", round(regression_jackknife$b1, 2), "SE(b1) =", round(regression_jackknife$se_b1, 2)))

# Extract jackknife estimates for histogram plotting
jack_b0 <- regression_jackknife$jack_b0
jack_b1 <- regression_jackknife$jack_b1

# Plot histograms of Jackknife b0 and b1
par(mfrow=c(1,2))  # Set up a 1x2 plotting area
hist(jack_b0, main="Jackknife b0", xlab="Intercept (b0)", col="lightblue", breaks=30)
hist(jack_b1, main="Jackknife b1", xlab="Slope (b1)", col="lightgreen", breaks=30)

```

3. Compare the estimates for Q1 and Q2 obtained from normal theory, bootstrap, and jackknife.

Ans:
1. Jackknife and Normal Theory produce identical mean estimates, suggesting that removing one observation at a time does not introduce noticeable bias in this dataset.Bootstrap estimates are slightly lower, likely due to the additional variability introduced by resampling with replacement.

2. Bootstrap estimates for both Fish_Mean and Copepod_Mean are slightly lower, likely due to variations introduced by the resampling process.

3. Normal Theory and Jackknife produce the same standard error, indicating that the Jackknife method did not alter the standard error estimation in this case. Different standard error from the Bootstrap method shows that this approach suggests greater uncertainty in the data, which may affect confidence levels.

calculate the SE of intercept(b0) and slope(b1) of regression using normal theory 
```{r}
# Linear Regression and SE for b0 and b1
model <- lm(fish ~ copepod)
b0 <- coef(model)[1]
b1 <- coef(model)[2]
summary_model <- summary(model)
se_b0_normal <- summary_model$coefficients["(Intercept)", "Std. Error"]
se_b1_normal <- summary_model$coefficients["copepod", "Std. Error"]

print("Normal Theory estimates for Regression Coefficients:")
print(paste("b0: SE =", se_b0_normal))
print(paste("b1: SE =", se_b1_normal))
```

create dataframe for comparison
```{r}

comparison_table <- data.frame(
  Method = c("Normal Theory", "Bootstrap", "Jackknife"),
  Fish_Mean = c(mean_fish, mean(boot_means_fish), fish_jackknife$mean),
  Fish_SE = c(se_fish, se_boot_fish, fish_jackknife$se),
  Copepod_Mean = c(mean_copepod, mean(boot_means_copepod), copepod_jackknife$mean),
  Copepod_SE = c(se_copepod, se_boot_copepod, copepod_jackknife$se),
  b0 = c(b0, mean(boot_b0), regression_jackknife$b0),
  SE_b0 = c(se_b0_normal, se_boot_b0, regression_jackknife$se_b0), 
  b1 = c(b1, mean(boot_b1), regression_jackknife$b1),
  SE_b1 = c(se_b1_normal, se_boot_b1, regression_jackknife$se_b1) 
)

print(comparison_table)

```




