---
title: "HW4"
author: "Annie"
date: "2025-03-15"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Compute the regression coefficients for fish = b0+b1*copepod and using bootstrap method to calculate the 95% confidence limits of b1 and test whether b1 is significantly different from 0 (with bootstrap 1000 times). 
Please calculate the CL using three methods: percentile, BC and BCa methods.
```{r}
library(readxl)

# Define the file path (modify this to the actual location of your file)
#file_path <- "C:/Users/User/Desktop/Master 1.2/R/HW/enviANDdensity.xls"

# Read the Excel file
data <- read_excel("C:/Users/User/Desktop/Master 1.2/R/HW/enviANDdensity.xls")

# Extract the required columns and rename them for convenience
fish <- data[["FishDensity (ind./1000m3)"]]
copepod <- data[["CopepodDensity ind./m3"]]
```



```{r}
model <- lm(fish ~ copepod)
coef_original <- coef(model) # Extracts the regression coefficients from the model and stores them in coef_original.

# Bootstrap regression coefficients
B <- 999  # Perform 999 bootstrap iterations (original data + 999 resamples)
b1_bootstrap <- numeric(B + 1)  # Add one more space for the original data's coefficient
b1_bootstrap[1] <- coef_original[2]  # Store the original model's b1 (slope) in the first position

n <- length(fish)

for (i in 1:B) {
  idx <- floor(runif(n, min = 1, max = n + 1))  # Manually implement random sampling
  fish_resample <- fish[idx]  # Sample from fish and copepod to generate a new bootstrap sample.
  copepod_resample <- copepod[idx]
  model_resample <- lm(fish_resample ~ copepod_resample)
  b1_bootstrap[i + 1] <- coef(model_resample)[2]  
  #Store bootstrap b1 coefficient, starting from the second position.
  #Extract the regression coefficient b1 (slope) and store it in b1_bootstrap[i + 1].
  #This gives us B different estimates of b1, which are used to calculate the confidence interval.
}

```

Compute confidence intervals (Percentile method)
```{r}
CI_percentile <- quantile(b1_bootstrap, probs = c(0.025, 0.975))
print(CI_percentile)
```

Compute bias-corrected (BC) confidence intervals for b1 using the bootstrap method
```{r}
# Step 1: Calculate the bias correction factor z0
z0 <- qnorm(mean(b1_bootstrap < coef_original[2]))  
# b1_bootstrap < coef_original[2]` if bootstrap estimate is less than the original b1, creates a logical vector (TRUE/FALSE)  
# mean(b1_bootstrap < coef_original[2])` calculates the proportion of bootstrap estimates less than the original b1  
# qnorm(): converts this proportion into a z-score (z0), representing the bias correction factor  

# Step 2: Compute adjusted quantiles for confidence intervals  
alpha1 <- pnorm(2 * z0 + qnorm(0.025))  
# qnorm(0.025)` gives the standard normal quantile for the lower 2.5% percentile (-1.96)  
# 2 * z0 + qnorm(0.025)` adjusts the percentile based on bias correction  
# pnorm(...)` converts this adjusted z-score back to a cumulative probability

alpha2 <- pnorm(2 * z0 + qnorm(0.975))  
# qnorm(0.975)` gives the standard normal quantile for the upper 97.5% percentile (+1.96)  
# 2 * z0 + qnorm(0.975)` adjusts the upper percentile based on bias correction  
# pnorm(...)` converts this adjusted z-score back to a cumulative probability  

# Step 3: Compute the bias-corrected (BC) confidence interval  
CI_BC <- quantile(b1_bootstrap, probs = c(alpha1, alpha2))  
# quantile(b1_bootstrap, probs = c(alpha1, alpha2))`  
# extracts the bootstrap estimates corresponding to the adjusted percentiles,  
# providing the bias-corrected 95% confidence interval for b1.

print(CI_BC)
```

BCa method (requires computing acceleration value a, detailed derivation omitted) (formula at P15)
```{r}
a_hat <- sum((b1_bootstrap - mean(b1_bootstrap))^3) / (6 * sum((b1_bootstrap - mean(b1_bootstrap))^2)^(3/2))
#calculation of the skewness correction factor (a_hat), reflects the degree of skewness in the data distribution.
alpha1_BCa <- pnorm(z0 + (z0 + qnorm(0.025)) / (1 - a_hat * (z0 + qnorm(0.025))))#The probability of the lower bound
alpha2_BCa <- pnorm(z0 + (z0 + qnorm(0.975)) / (1 - a_hat * (z0 + qnorm(0.975))))#The probability of the upper bound
CI_BCa <- quantile(b1_bootstrap, probs = c(alpha1_BCa, alpha2_BCa))#Calculate BCa

# Test whether b1 is significantly different from 0
p_value <- mean(b1_bootstrap < 0) * 2  # Two-tailed test

# Display results
list(
  #coef_original = coef_original,
  CI_percentile = CI_percentile,
  CI_BC = CI_BC,
  CI_BCa = CI_BCa,
  p_value = p_value
)

```
2. Bootstrap test whether significant difference exists between the density of Oncaea Venusta and Canthocalanus pauper, using and BCa. (Assume each station is independent and use all 34 stations.)


Extract density
```{r}
copepod_data <- read.table("copepod_composition.txt", header = TRUE)
cop_density <- read.table("cop_density.txt", header = TRUE)
species_list <- read.table("copepodSPlist.txt", fill=T, sep = "\n")

row.names(copepod_data) <- species_list$V1

Oncaea <- copepod_data ["Oncaea venusta", ]/100
Pauper <- copepod_data ["Canthocalanus pauper", ]/100

Oncaea_new <- matrix(0, nrow = 1 , ncol = length(species_list$V1))
row.names(Oncaea_new) <- "Oncaea venusta"
for( i in 1: length(cop_density$X1119)) {
  Oncaea_new[ ,i] <- Oncaea[, i]*cop_density[i,]
  }
  
Pauper_new <- matrix(0, nrow = 1 , ncol = length(species_list$V1))
row.names(Pauper_new) <- "Canthocalanus pauper"
for( i in 1:length(cop_density$X1119)) {
  Pauper_new[ ,i] <- Pauper[, i]*cop_density[i,]
}

```

```{r}
# Define the bootstrap function
bootstrap_stat <- function(data, indices) {
  density1_sampled <- data$density1[indices]
  density2_sampled <- data$density2[indices]
  return(mean(density1_sampled - density2_sampled))
}

# Set Bootstrap parameters
B <- 999  
n <- length(Oncaea_new[1, ])  

# Get original density data
density1 <- as.numeric(Oncaea_new["Oncaea venusta", ])  
density2 <- as.numeric(Pauper_new["Canthocalanus pauper", ])  

# **Store data in a DataFrame for easy access**
boot_data <- data.frame(density1 = density1, density2 = density2)

# **Perform Bootstrap**
library(boot)
boot_results <- boot(data = boot_data, statistic = bootstrap_stat, R = B)

# **Compute BCa confidence interval**
bca_result <- boot.ci(boot_results, type = "bca")

# **Print the results**
print(paste("Original mean difference:", mean(density1 - density2)))
print(paste("Bootstrap mean difference:", mean(boot_results$t)))
print(paste("BCa confidence interval:", bca_result$bca[4], "to", bca_result$bca[5]))


```

