---
title: "HW5"
author: "Annie"
date: "2025-03-20"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Compute the regression coefficients for fish = b0+b1*copepod and use randomization method to generate null distribution of b1 and test whether b1 is significantly different from null (with randomization 5000 times). Report your p-value. 
```{r}
library(readxl)

# Read the Excel file
data <- read_excel("C:/Users/User/Desktop/Master 1.2/R/HW/enviANDdensity.xls")

# Extract the required columns and rename them for convenience
fish <- data[["FishDensity (ind./1000m3)"]]
copepod <- data[["CopepodDensity ind./m3"]]
```

```{r}
# Compute original regression coefficient
lm_original <- lm(fish ~ copepod)
b1_original <- coef(lm_original)[2]
n <- length(fish)

# Define shuffle function using order(runif(n))
fisher_yates_shuffle <- function(n) {
  return(order(runif(n)))  # Generate a random order of indices
}

# Define randomization function
randomization_test <- function(data, indices) {
  fish_shuffled <- data$fish[indices]  # Shuffle fish values using indices
  model <- lm(fish_shuffled ~ data$copepod)
  return(coef(model)[2])
}

# Prepare data frame for randomization
set.seed(123)  # Set random seed for reproducibility
data_for_perm <- data.frame(fish = fish, copepod = copepod)

# Perform 5000 randomizations
perm_results <- replicate(5000, {
  shuffled_indices <- fisher_yates_shuffle(length(fish))  # Use order(runif(n))
  randomization_test(data_for_perm, shuffled_indices)
})

# Compute p-value (two-tailed test)
p_value <- mean(abs(perm_results) >= abs(b1_original))

# Output results
print(paste("Original regression coefficient b1:", b1_original))
if (p_value < 0.05) {
  print("p-value: < 0.05")
} else {
  print(paste("p-value:", p_value))
}

```

2. Randomization test whether significant difference exists between the density of Oncaea Venusta and Canthocalanus pauper. (Assume all data are independent and use all 34 stations.) Report your pvalue. 
NOTE: no built-in randomization routines are allowed!

Extract density
```{r}
copepod_data <- read.table("copepod_composition.txt", header = TRUE)
cop_density <- read.table("cop_density.txt", header = TRUE)
species_list <- read.table("copepodSPlist.txt", fill=T, sep = "\n")

row.names(copepod_data) <- species_list$V1

Oncaea <- copepod_data ["Oncaea venusta", ]/100
Pauper <- copepod_data ["Canthocalanus pauper", ]/100

Oncaea_new <- matrix(0, nrow = 1 , ncol = length(species_list$V1))
row.names(Oncaea_new) <- "Oncaea venusta"
for( i in 1: length(cop_density$X1119)) {
  Oncaea_new[ ,i] <- Oncaea[, i]*cop_density[i,]
  }
  
Pauper_new <- matrix(0, nrow = 1 , ncol = length(species_list$V1))
row.names(Pauper_new) <- "Canthocalanus pauper"
for( i in 1:length(cop_density$X1119)) {
  Pauper_new[ ,i] <- Pauper[, i]*cop_density[i,]
}

```

```{r}
set.seed(123)  # Set the seed for reproducibility

# Calculate the observed paired difference
# We calculate the density difference between Oncaea and Pauper at each station and then take the mean
observed_diff <- mean(Oncaea_new - Pauper_new)

# Step 2: Number of randomizations
n_randomizations <- 5000
random_diffs <- numeric(n_randomizations)

# Step 3: Perform the randomization test
for (i in 1:n_randomizations) {
  
  # Step 3.1: Combine the densities of Oncaea and Pauper into one vector
  combined_densities <- c(Oncaea_new, Pauper_new)
  
  # Step 3.2: Generate random numbers and order them
  random_numbers <- runif(length(combined_densities))  # Generate 362 numbers that ranged between 0&1
  random_index <- order(random_numbers)  # Order random numbers from big to small
  
  # Step 3.3: Assign the shuffled densities back to Oncaea and Pauper
  shuffled_Oncaea <- combined_densities[random_index[1:length(Oncaea_new)]] #Select the first length(Oncaea_new)(182!) elements from the random_index vector, so that the corresponding values at these positions can be chosen from combined_densities.
  shuffled_Pauper <- combined_densities[random_index[(length(Oncaea_new) + 1):length(combined_densities)]]#select number 183~362 elements from the random_index vector
  
  # Step 3.4: Calculate the difference after randomization
  random_diffs[i] <- mean(shuffled_Oncaea - shuffled_Pauper)
}

# Step 4: Calculate the p-value
# The p-value is the proportion of randomized differences that are as extreme or more extreme than the observed difference
p_value <- mean(abs(random_diffs) >= abs(observed_diff))

# Output the p-value
p_value

```
把所有密度混在一起，然後再分給Oncaea和Pauper，然後再看新的Oncaea還有Pauper有沒有差很多已決定whether significant difference exists between the density of Oncaea Venusta and Canthocalanus pauper!

