---
title: "HW7"
author: "Annie"
date: "2025-04-17"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Use the dominant copepod species data (from HW1).

Perform cluster analysis of stations based on percent composition data of the dominant species, and tell your story about these copepod data.

Tasks:

Compare the results based on different distance measures.

Compare the results based on different cluster algorithms.

Determine the final number of clusters and describe the differences among them.

You are allowed to use any built-in functions in Matlab or R, such as kmeans.m, clusterdata.m, etc.
```{r}
copepod_data <- read.table("copepod_composition.txt", header = TRUE)
cop_density <- read.table("cop_density.txt", header = TRUE)
species_list <- read.table("copepodSPlist.txt", fill=T, sep = "\n")

row.names(copepod_data) <- species_list$V1


# Prepare an empty matrix to store density data for all species
all_species_density <- matrix(0, nrow = nrow(copepod_data), ncol = ncol(copepod_data))
row.names(all_species_density) <- row.names(copepod_data)
colnames(all_species_density) <- colnames(copepod_data)

# Loop through each species (row) and each station/sample (column) to calculate density
for (i in 1:nrow(copepod_data)) {
  for (j in 1:ncol(copepod_data)) {
    # Species proportion × corresponding total density at that station
    all_species_density[i, j] <- copepod_data[i, j] / 100 * cop_density[j, 1]
  }
}

# Identify dominant species
# A dominant species is defined as one that makes up ≥5% of the total composition at any cruise-station
threshold <- 5  # Threshold in percent

# Find species that exceed the threshold at least once across all stations
dominant_species <- rownames(copepod_data)[apply(copepod_data, 1, function(x) any(x >= threshold))]

# Display the dominant species
print(dominant_species)


```
Compare the results based on different distance measures.
```{r}
#select dominant species only
dominant_comp <- copepod_data[rownames(copepod_data) %in% dominant_species, ] 

library(vegan)  # for Bray-Curtis
library(cluster)  # for clustering visualization
library(factoextra)  # for visualizing clusters

# Transpose: clustering is based on stations (columns), not species (rows)
dominant_t <- t(dominant_comp)

# Standardize data (optional, based on range)
dominant_scaled <- scale(dominant_t)

# Different distance measures
dist_euclidean <- dist(dominant_scaled, method = "euclidean")
dist_bray <- vegdist(dominant_t, method = "bray")

# Hierarchical clustering
hc_euclidean <- hclust(dist_euclidean, method = "ward.D2")
hc_bray <- hclust(dist_bray, method = "ward.D2")

# Plot dendrograms
par(mar = c(5, 4, 4, 2))
plot(hc_euclidean, main = "Hierarchical Clustering (Euclidean)", xlab = "", sub = "", cex=0.7)
plot(hc_bray, main = "Hierarchical Clustering (Bray-Curtis)", xlab = "", sub = "", cex=0.7)

```
Compare the results based on different cluster algorithms.
```{r}
# Determine number of clusters with Elbow method
fviz_nbclust(dominant_scaled, kmeans, method = "wss") + 
  labs(subtitle = "Elbow Method for K-means")

# Perform K-means with optimal k (say k = 3)
set.seed(123)
km_res <- kmeans(dominant_scaled, centers = 3, nstart = 25)


# Plot clustering result
cluster_plot <- fviz_cluster(km_res, data = dominant_scaled, geom = "point",
                             main = "K-means Clustering (k = 3)")

# Add location name
library(ggplot2)
cluster_plot + 
  geom_text(aes(label = rownames(dominant_scaled)), 
            vjust = -0.8, size = 3)

```
Determine the final number of clusters and describe the differences among them.
```{r}
# Add cluster labels to data
cluster_groups <- km_res$cluster

# Check species composition differences by cluster
dominant_with_cluster <- cbind(dominant_t, cluster = factor(cluster_groups))

# Convert the matrix to a data frame
dominant_df <- as.data.frame(dominant_with_cluster)

# Add the cluster result from K-means as a new column
dominant_df$cluster <- factor(km_res$cluster)

# Calculate the average species composition for each cluster
cluster_summary <- aggregate(. ~ cluster, data = dominant_df, FUN = mean)


print(cluster_summary)

```
Cluster 1 appears to be dominated by certain larger copepod species like Acartia pacifica and Calanus sinicus, along with some Paracalanus species.
Cluster 2 shows a strong presence of Paracalanus pavus and several smaller cyclopoid copepods like Oithona similis and Corycaeus (Ditrichocorycaeus) affinis.
Cluster 3 is characterized by high abundances of other small cyclopoid copepods such as Parvocalanus crassirostris, Oithona species, and Euterpina acutifrons, as well as Canthocalanus pauper.
